\textbf{Abstract.} Smart Cities are cities that employ sensors, actors, communication protocols and home servers to increase the efficiency of city services and processes, including water and power supply, waste disposal, police, fire/rescue, hospitals, and schools. Information and Communication Technology (ICT) helps optimizing the processes, while the Internet of Things (IoT) provides the platform for managing a multitude of small sensor and actor devices. A smart city promises not only a more efficient management of resources, but also an increase in the quality of services provided, while still remaining cost-effective. 

From a technical point of view, a smart city like any other "Smart X" systems can be seen as a large, sensor-based, distributed information system with data collection, data processing and data support components. Nowadays, smart systems and software architectures are reaching new levels of complexity, necessitating appropriate engineering methodologies. Model-Driven Software Engi­neer­ing provides the required foundations for formally define generic architectures and development "Smart X" software support on the top. To cope with ever growing complexity of such software architectures, model-driven approaches are extensively applied to model IoT technologies as well as developing and maintaining software and systems platforms for "Smart X" technologies. As the models of these architectures and platforms become huge and complex over time, model-driven collaborative development is necessary to cope with development and evolution challenges arising from the IoT applications. This research proposal aims at (1) studying the state of the art in model-driven IoT and collaborative model-driven IoT, (2) extended research in model-driven IoT focusing on the "Smart X", (3) collaborative model-driven IoT for developing software architectures and platforms for IoT-based systems. As the proof of concepts, it further focuses on applying model-driven collaborative development and maintenance, new upcoming trends in collaborative modeling of IoT, and its adoption in developing smart city architectures.