Smart cities combine innovative Information and Communication Technologies (ICT) to improve urban services aiming at overcoming social, economic and environmental changes. 

The internet of the present day is enriched with a huge amount of aforementioned various devices and micro services. These devices and services operate for different purposes, yet work together to achieve common goals. The various types of sensors, actuators and devices are being developed to provide a range of services. These sensors and devices have created a new technological trend today. This trend has been named the "Internet of Things" \cite{WeberRomana2010}, "Network of Things" \cite{Gubbi+2013} or "Web of Things" \cite{Guinard+2010}. 

In the framework of IoT, more and more devices are equipped with network connectivity to autonomously provide "smarter" services, forming the Internet of Things (IoT). Applications are wide-ranging, and have variously been termed "Smart X", including Smart Homes, Smart Factories (Industry 4.0), Smart Government, Smart City, Smart Grid, Smart Traffic Control, and many more.

Smart Cities are cities that employ sensors and actors to increase the efficiency of city services and processes, including environmental protection/sustainability, energy efficiency, mobility, health care, and safety/security. ICT helps to optimize the processes, while the Internet of Things (IoT) provides the platform for managing a multitude of small sensor, actor devices, communication protocols and home servers.

\textbf{Internet of Things (IoT)} \cite{WeberRomana2010} is the basic concept for developing the network of things equipped with built-in technologies for interaction with each other or with the environment. The concept was formulated in 1999 as an understanding of the prospects for the widespread use of radio frequency identification means for the interaction of physical objects with each other and with the external environment. 
\begin{definition}
\textbf{Internet of Things}\label{definition:internetOfThings}\\
The Internet of Things is about installing sensors (RFID, IR, GPS, laser scanners, etc.) for everything, and connecting them to the internet through specific protocols for information exchange and communications, in order to achieve intelligent recognition, location, tracking, monitoring and management [reference].
\end{definition}

\textbf{Smart City}. The concept of integrating several ICT and IoT solutions for the management of urban property. The purpose of creating a "smart city" is to improve the quality of life with the help of computer technologies to improve the efficiency of services and meet the needs of people. ICT allows the city government to directly interact with communities and urban infrastructure, and monitor what is happening in the city, how the city is developing, and what can improve the quality of life. Through the use of sensors integrated in real time, the accumulated data from urban residents and devices are processed and analyzed. 

The concept of smart city arises from the need to manage, automate, optimize and explore all aspects of a city that could be improved and optimized by information technologies. The software paradigm IoT, being a core concept behind smart cities, is largely perceived as a collection of interconnected "things" within smart cities.

The IoT-based smart city applications are realized by interconnected systems of heterogeneous hardware, software, and embedded systems: these cyber-physical systems introduce new levels of complexity, requiring appropriate engineering methodologies to support formally rigorous software and systems development \cite{Kuryazov+2018}. Model-Driven Software Engineering (MDE) provides fitting foundations and is considered as an enabling technology for advancing Smart X applications.

A number of model-driven reference architectures \cite{KateuleWinter2018} and models \cite{Yin+2015} are introduced for smart city development, so far. These architectures and models contemplate smart city architectures as a blueprint which provides an appropriate level of abstraction for the development process of smart cities. Model-driven reference architectures and models are used to represent and define different development aspects of smart city architectures gathering different views, viewpoints, soft- ware components, communication protocols, sensors, activators, etc. in a single, huge architecture. For instance, several model-driven approaches are investigated in [12] utilizing different modeling language profiles (e.g. standard UML profiles [3]) in development of smart city architectures. The standard UML profiles is also perfectly aligned in [15] for developing smart city architectures based on Internet of Things.

MDE has a much wider set of modeling techniques and a much more detailed separation of different views and concerns [7]. For instance, components have well-defined interfaces and ports. MDE benefits from its capabilities to model different, but integrated views and behavior of systems which are eventually made executable for different platforms. The basic idea of the smart city vision is the pervasive presence of a variety of things or objects such as sensors, actuators, mobile phones, etc. which are able to interact with each other and cooperate with their pairs to reach common goals [1][5]. This section shortly reviews MDE approaches and methodologies for developing smart city applications that the collaborative MDE infrastructure can be applied to.

As a software engineering paradigm, MDE is the modern day style of software and system development which supports well-suited abstraction concepts to development activities. Model-driven engineering intends to improve the productivity of the design and develop­ment, maintenance activities, and communication among various actors and stakeholders of a system. In MDE, software models (e.g. in UML [reference]) which also comprise source code are the central key artifacts. They are well-suited for designing, developing and producing large-scale software projects. Software models are the documentation and implementation of software systems being developed and evolved [reference]. MDE brings several main benefits such as a productivity boost and , models become a single point of truth., and theyModels are reusable and automatically kept in sync with the code they represent [reference]. Software models (e.g. in UML [reference]) are the key artifacts in MDE activities. They are well-suited for designing, developing and producing large-scale software projects. Software models are the documentation and implementation of software systems being developed and evolved [reference].

Models are constantly changed during their development and evolutionary life-cycle by various developers and experts. They are constantly evolved and maintained undergoing diverse changes such as extensions, corrections, optimization, adaptations and other improvements. All development and maintenance activities contribute to the evolution of software models resulting in several subsequent revisions. During software evolution, models become large and complex raising a need for collaboration of several developers, designers and stakeholders (i.e. collaborators) on shared models i.e. \textit{Collaborative modeling}.

A meta-model generic, modeling tool generic and flexible collaborative modeling infrastructure was introduced in [4]. The associated technical support also focuses on providing a catalog of supplementary services which are underlying services for developing collaborative modeling applications.

This research proposal is structured as follows: Section~\ref{sec:motivation} motivates the research field defining its contributions and novelty. Section~\ref{sec:stateofart} investigates existing model-driven IoT approaches and collaborative modeling approaches for IoT. The main research objectives of this proposal are sketched in Section~\ref{sec:researchObjectives}. This paper ends up in Section~\ref{sec:conclusion} by drawing some conclusions.